Separating Use and Reuse to Improve Both

### 2017-04-03 FACS/Extensions.tex
added example in intro

\section{More composition operators}
\use is amazing, elegant and simple, but our system can be easily enriched with more 
operators: while most approaches in literature presents a fixed set of operators, we do not need such restriction.
We just need to be sure that every newly added operator respects the following criteria:

@@

\paragraph*{Application on the expression problem}${}_{}$\\*

With redirect+rename we can have the general operator propagator

operation:{//for sum and lit, easy to extends as before
  T:{}
  Exp:{interface method Exp op(T x)}}
  Sum:Use sum,{ extends Exp sum,expAdd1,{
    method Exp op(T x)
      Sum.of(left.op(x),right.op(x))  }
  Lit:Use lit,{
    method Exp op(T x)  this
  }
Now, to have my addN I can
opAddn: Use
  operation[redirect T to Int]
    [rename Exp.op(x) to addN(x)][restrict Lit.op(x)], {
  Lit:{method Exp addN(Int x) Lit.of(inner())+x}
  }

Full power of redirect

serviceCombinator:{ S:{implements Service
  method R report()
  }

  R:{method R combine(R that)
    class method R empty()}

@@

  class R method(ListS ss){//here we use extended java like syntax
    R r=R.empty()
    for(S s in ss){
      s.performService();
      r=r.combine(s.report())
      }
    return r;
  }
}
PaintingService:serviceCombinator[redirect S to Painting.Service]

To encode the former generic code in java you need to write
the following headeche inducing interfaces for RService and Report (and still we miss the empty method)
and require that the services you want to serve implement those.
  interface Service{ void performService();}
  interface Report<R extends Report<R>>{R combine(R that);}
  interface RService<R extends Report<R>> extends Service{ R report();}

>>>>

\section{More composition operators}
\use\ is amazing, elegant and simple, but our system can be easily enriched with more 
operators: while most approaches in literature presents a fixed set of operators, we do not need such restriction.
We just need to be sure that every newly added operator respects the following criteria:

@@

\paragraph*{Application on the expression problem}${}_{}$\\*
With redirect, rename and restrict we can have the general operator propagator
\begin{lstlisting}
operation:{//for sum and lit, easy to extends as before
  T:{}
  Exp:{interface method Exp op(T x)}}
  Sum:Use sum,{ extends Exp sum,expAdd1,{
    method Exp op(T x)
      Sum.of(left.op(x),right.op(x))  }
  Lit:Use lit,{
    method Exp op(T x)  this
  }
\end{lstlisting}

Now, to have my addN I can

\begin{lstlisting}
opAddn: Use
  operation[redirect T to Int]
    [rename Exp.op(x) to addN(x)][restrict Lit.op(x)], {
  Lit:{method Exp addN(Int x) Lit.of(inner())+x}
  }
\end{lstlisting}

\paragraph*{Full power of redirect}${}_{}$\\*
An expressive form of Redirect can be multiple, that is, can redirect may interdependent classes at the same time.
We show an example where a specific kind of \Q@Service@ can produce a \Q@Report@, and 
\Q@Report@s can be combined together.
The goal is to execute a list of such services and produce a collated report.
This example also show how to propagate generics:

\begin{lstlisting}
Service:{interface method Void performService()}
serviceCombinator:{
  S:{implements Service method R report()  }

  R:{method R combine(R that)   class method R empty() }

@@

  class method R doAll(ListS ss){//here we use extended java like syntax
    R r=R.empty()
    for(S s in ss){
      s.performService();
      r=r.combine(s.report())
      }
    return r;
  }
}
PaintingService:serviceCombinator[redirect S to PaintingService]
PaintingService:{... method PaintingReport report()..}
PaintingReport:{..}
\end{lstlisting}

The flattened version of \Q@PaintingService@ would look like:
\begin{lstlisting}
PaintingService:{
  ListS:/*the expansion of list[redirect Elem to PaintingService]*/

  class method PaintingReport doAll(ListS ss){
    PaintingReport r=PaintingReport.empty()
    for(PaintingService s in ss){
      s.performService();
      r=r.combine(s.report())
      }
    return r;
  }
}
\end{lstlisting}
Where you can note how redirect figured out \Q@R=PaintingReport@ by comparing the structural shape of
classes \Q@PaintingService@ and \Q@S@.

To encode the former generic code in java you need to write
the following headeche inducing interfaces for RService and Report.
and require that the services you want to serve implement those.
\begin{lstlisting}
interface Service{ void performService();}
interface Report<R extends Report<R>>{R combine(R that);}
interface RService<R extends Report<R>> extends Service{ R report();}
\end{lstlisting}
Note how we still can not encode the method \Q@empty@.

### 2017-04-03 FACS/Intro.tex
added example in intro

Class \lstinline{B} now have a method \lstinline{ma} inherited from \lstinline{A}.
This method passes \lstinline{this} as \lstinline{A}; we can see an invocation of \lstinline{ma}
inside \lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.
The execution will eventually call \lstinline{Util.m} with a \lstinline{B} instance.
This can be correct only if \lstinline{B} is a subtype of \lstinline{A}.
Thus, every OO language with the minimal features 
exposed in this example (using this, extends, method call)
is forced to accept that subclassing implies subtyping.
Note how this is true also in C++, where is possible to
"extends privately". Such is a limitation of the visibility of
subtyping but not over subtyping itself, and the former example
would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.
\marco{Cite some work of bruceTHisType and show how he also fails.}
\bruno{I was expecting some text/argument here on why subclassing
  implies subtyping is bad. Perhaps another example where subclassing 
must not imply subtyping to type-check.}

@@

>>>>

Class \lstinline{B} now have a method \lstinline{ma} inherited from \lstinline{A}.
This method passes \lstinline{this} as \lstinline{A}; we can see an invocation of \lstinline{ma}
inside \lstinline{mb}, where \lstinline{this} is of type \lstinline{B}.
The execution will eventually call \lstinline{Util.m} with a \lstinline{B} instance.
This can be correct only if \lstinline{B} is a subtype of \lstinline{A}.
Thus, every OO language with the minimal features 
exposed in this example (using this, extends, method call)
is forced to accept that subclassing implies subtyping.
Note how this is true also in C++, where is possible to
"extends privately". Such is a limitation of the visibility of
subtyping but not over subtyping itself, and the former example
would be accepted by C++ even if \lstinline{B} was to "privately extends" \lstinline{A}.
\marco{Cite some work of bruceTHisType and show how he also fails.}
\bruno{I was expecting some text/argument here on why subclassing
  implies subtyping is bad. Perhaps another example where subclassing 
must not imply subtyping to type-check.}
An iconic example on why connecting inheritance/code reuse and subtpying is problematic is
provided by the historic[]:
A reasonable implementation for a \Q@Set@ may be easy to extend into a \Q@Bag@ keeping tracks of how many times an element occurs.
We would just add some state and override a couple of methods.

However, our subclassing would break LSP[]: not all bags are sets!
Of course, one could retroactivelly fix this problem by introducing \Q@AbstractSetOrBag@
and making both \Q@Bag@ and \Q@Set@ inherit from it.
This looks innatural, since \Q@Set@ would extend it without adding anything,
and we would be suprtised to find a use of the type \Q@AbstractSetOrBag@.
Worst, if we was to constantly apply this mentalty, we would introduce a very high number
of abstract classes that are not supposed to be used as types, and that will clutter the 
public interface of our classes and our code project as a whole.

@@

\subsection{List of contribution}
\marco{text to improve, but is nice to keep a list}
\begin{itemize}
\item simple approach to fully separate inheritance and subtyping
\item easy to grasp model of compilation were typing is late but not too late (lazy?)
\item clean and elegant handling of state in trat/module composition language
\item extension with nested class is natural and powefull
\item more operators can be accomodated without changing the general model
\end{itemize}
